---
title: "Extending GLMs"
output: pdf_document
---

\renewcommand{\vec}[1]{\mathbf{#1}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 4, fig.width = 6, fig.align = 'center')
library(tidyverse) 
library(rstanarm)
library(arm)
library(gridExtra)
library(rstan)
library(brms)
library(extraDistr)
set.seed(02012021)
```

#### Heteroscedastic Models


\vfill

```{r}
n <- 500
x <- runif(n,0,10)
beta0 <- 1
beta1 <- 2
y <- rnorm(n, mean = beta0 + x * beta1, sd = x *sqrt(2))

tibble(y = y, x = x) %>% ggplot(aes(y = y, x = x)) +
  geom_point() + theme_bw() + 
  geom_smooth(formula = 'y~x', method ='lm')
```
\vfill
\newpage

Stan code can be written to estimate the variance as a function of x.

```{stan output.var='text_only', eval = F}
data {
  int<lower=0> N;
  vector[N] y;
  vector[N] x;
}






```


\vfill
```{r, results = 'hide', message = F, error = F, warning = F}
reg_ncv <- stan("heteroskedastic_regression.stan", data=list(N = n, y=y, x = x), refresh = 0)
```

\vfill

```{r}
print(reg_ncv)
```

\newpage

## Mixture Models

Sometimes a single probability distribution isn't sufficient to model an outcome of interest. 

\vfill

\vfill

\vfill

\vfill

\vfill




\vfill



\newpage

This model could be coded in stan or consider using the `brms` package (bayesian regresion models in stan)

\vfill

```{r, message = F, results = 'hide', warning = F}
zero_prob <- .33
n <- 1000
indicator <- rbinom(n,1,zero_prob)
counts <- tibble(counts = rpois(n, lambda = 15) * (1 - indicator))

zip <- brm(counts ~ 1, data = counts, 
           family = zero_inflated_poisson, refresh = 0)
# zip <- brm(counts ~ 1, data = counts, 
#family = zero_inflated_poisson, refresh = 0, save_model = 'zip')
```

\vfill

```{r}
print(zip)
```


\newpage



\vfill

\vfill

```{r}
trunc_pois <- tibble(y = rtpois(n, 5, a = 0, b = Inf))

trunc_pois %>% ggplot(aes(x = y)) + 
  geom_bar() + theme_bw() + xlim(0, NA)
```

\vfill


```{r, results = 'hide', warning =F, message = F}
truncated_pois <- brm(y ~ 1, data = trunc_pois, family = hurdle_poisson, refresh = 0)
```

\newpage

```{r, warning = F, message = F}
print(truncated_pois)
```

\vfill

\newpage


```{r, results = 'hide'}
hurdle_pois <- trunc_pois %>% bind_rows(tibble(y = rep(0, n)))
hurdle_poisson <- brm(y ~ 1, data = hurdle_pois, family = hurdle_poisson, refresh = 0)
```

```{r}
print(hurdle_poisson)
```

#### Discrete / Continuous


```{r}
zero_prob <- .33
n <- 1000
indicator <- rbinom(n,1,zero_prob)
counts <- tibble(counts = rlnorm(n, meanlog = log(5)) * (1 - indicator))

head(counts)

counts %>% ggplot(aes(x = counts)) + geom_histogram(bins = 40)
```


