---
title: "Ordinal Regression + Compound Models"
output: pdf_document
---

\renewcommand{\vec}[1]{\mathbf{#1}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 4, fig.width = 6, fig.align = 'center')
library(tidyverse) 
library(rstanarm)
library(arm)
library(gridExtra)
library(brms)
library(optpart)
set.seed(01252021)
library(mvabund)
data("shoshveg")
data("shoshsite")
```

### Categorical Data

*Categorical data comes in two different flavors:*
\vfill
*1. Ordinal: Likert scale responses*
\vfill
*2. Unordered: favorite winter Montana activity (skiing, sleeping, doing Statistics)*

\vfill

### Models for Ordinal Data


Recall the latent formulation for probit model


```{r, echo = F}
plot(seq(-2.5,2.5,.1), dnorm(seq(-2.5,2.5,.1)), type = 'l', axes = F , ylab = '', xlab ='')
```

\vfill

*Let $y$ be a vector categorical response*

\begin{align}
y &\sim Multinomial(\pi_1, ..., \pi_k)\\
\end{align}

*Using the latent data formulation, let $z$ be a continuous normal random variable with mean (XB) and variance 1. Then, for k = 3,*
\begin{equation*}
y_i = \begin{cases} 1 \text{ if }\; z_i < c_{1|2} \\
2 \text{ if }\;c_{2|3} \leq z_i \geq c_{1|2} \\
3 \text{ if }\; z_i > c_{2|3} \end{cases}
\end{equation*}

\vfill
\newpage

Then given $z_i \sim N(X_i \beta, 1),$ $\underline{\pi_i}$ can be calculated by integrating the area of the distribution in the appropriate cut points.

```{r, echo = F}
plot(seq(-2.5,2.5,.1), dnorm(seq(-2.5,2.5,.1)), type = 'l', axes = F , ylab = '', xlab ='')
```

For instance, suppose $X_i \beta = 2$ and $c_{1|2} = 0, c_{2|3} = 2$, then

$\pi_1 =$ `pnorm(0, mean = 2)` = (`r pnorm(0, mean = 2)`)

\vfill

$\pi_2 =$ `pnorm(2, mean = 2) - pnorm(0, mean = 2)` = (`r pnorm(2, mean = 2) - pnorm(0, mean = 2)`)

\vfill

$\pi_3 =$ `1 - pnorm(2, mean = 2)` = (`r 1 - pnorm(2, mean = 2)`)

\vfill

\newpage

#### Vegetation Coverage Class Data

Plant coverage is an important ecological indicator. However, estimating plant coverage can be difficult and is often summarized in an ordinal manner.

\vfill

```{r}
class <- shoshveg  %>% 
          mutate(class = case_when(
            ABILAS <= .5 ~ 1,
            ABILAS > .5 & ABILAS <= 3  ~ 2,
            TRUE ~3)) %>% 
          mutate(class = factor(class)) %>%
          dplyr::select(class) %>% pull()

plant_cover <- tibble(class = class, elevation = scale(shoshsite$elevation))
```

\vfill

```{r}
ord <- polr(class ~ elevation, data = plant_cover, method = 'probit')
summary(ord)
bayes_ord <- stan_polr(class ~ elevation, data = plant_cover, method = 'probit',
          prior = R2(0.25, 'mean'), refresh = 0)
print(bayes_ord, digits = 2)
```

