---
title: "Other Generalized Linear Models"
output: pdf_document
---

\renewcommand{\vec}[1]{\mathbf{#1}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 4, fig.width = 6, fig.align = 'center')
library(tidyverse) 
library(rstanarm)
library(arm)
library(gridExtra)
set.seed(01252021)
```

#### Logistic Binomial Model


For count data we have discussed Poisson and Negative-Binomial sampling models. 

\vfill


\vfill

```{r}
batting <- read_csv('http://math.montana.edu/ahoegh/teaching/stat491/data/BattingAverage.csv') %>% 
  mutate(NotHits = AtBats - Hits)
batting %>% sample_n(5)
```

\newpage


\vfill

```{r}
log_binom <- stan_glm(cbind(Hits, NotHits ) ~ PriPos - 1, 
              family = binomial(link = "logit"), data = batting, refresh = 0)

print(log_binom, digits = 2)
```

\vfill



Overdispersion can also occur with binomial data. Recall that the variance of binomial trials is $np(1-p)$. 

\vfill

\vfill


\newpage

#### Probit Model


\vfill


\vfill



\newpage

```{r}
probit_binom <- stan_glm(cbind(Hits, NotHits ) ~ PriPos - 1, 
              family = binomial(link = "probit"), data = batting, refresh = 0)
print(probit_binom)
plot(probit_binom)
```

\newpage

So what do the coefficients mean in this model?

\vfill

```{r}
log_binom$coefficients
probit_binom$coefficients
```

\vfill

```{r}
invlogit(log_binom$coefficients) * 1000
pnorm(probit_binom$coefficients) * 1000
```




