---
title: "Other Generalized Linear Models"
output: pdf_document
---

\renewcommand{\vec}[1]{\mathbf{#1}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 4, fig.width = 6, fig.align = 'center')
library(tidyverse) 
library(rstanarm)
library(arm)
library(gridExtra)
set.seed(01252021)
```

#### Logistic Binomial Model


For count data we have discussed Poisson and Negative-Binomial sampling models. *It is also possible to use a Binomial distribution, but know that the support of the response will not be countably infinite.*

\vfill

*A common example of binomial data would be free throw shooting for basketball players or batting data for baseball players.*

\vfill

```{r}
batting <- read_csv('http://math.montana.edu/ahoegh/teaching/stat491/data/BattingAverage.csv') %>% 
  mutate(NotHits = AtBats - Hits)
batting %>% sample_n(5)
```

\newpage

*The logistic-binomial framework is written as:*

\begin{align}
y_i \sim & Binomial(n_i, p_i), \\
logit(p_i) = & X_i \beta
\end{align}

\vfill

```{r}
log_binom <- stan_glm(cbind(Hits, NotHits ) ~ PriPos - 1, 
              family = binomial(link = "logit"), data = batting, refresh = 0)

print(log_binom, digits = 2)
```

\vfill



Overdispersion can also occur with binomial data. Recall that the variance of binomial trials is $np(1-p)$. *Then define the residuals as $z_i = \frac{y_i - \hat{y_i}}{sd(\hat{y_i})}$.*

\vfill
*Then the $z_i$ terms should be approximately iid N(0,1). A formal test for $\sum z_i^2$ using a $\chi^2$ distribution can be used to detect overdispersion.*

\vfill

*Often hierarchical models will solve some of these issues, otherwise an overdispersion model can be formulated with variance equal to $\omega np(1-p)$. See `brm` or write your own in stan.*

\newpage

#### Probit Model

*Consider an alternative link function for binary/binomial data.*

\begin{align}
y_i \sim & Binomial(n_i, p_i), \\
\Phi^{-1}(p_i) = & X_i \beta \\
p_i = & \Phi(X_i \beta),
\end{align}

*where $\Phi()$ is the cumulative distribution function for a standard normal random variable.*
\vfill

*This model is a latent data model, which are very common and useful in statistics. We assume there is an underlying continuous random variable that is mapped to a standard normal distribution.*

```{r, echo = F}
plot(seq(-4,4, by = .1), dnorm(seq(-4,4, by = .1), mean = -1), type = 'l', axes = F, xlab = '', ylab = '')
```

\vfill

where

\begin{equation*}
y_i = \begin{cases} 1 \text{ if }\; z_i > 0 \\
0  \text{ if }\; z_i < 0 \end{cases}
\end{equation*}

\begin{align*}
z_i = & X_i \beta + \epsilon\\
\epsilon \sim & N(0,1)
\end{align*}

Note $\epsilon \sim  N(0,1)$ is a necessary constraint for this model.

\newpage

```{r}
probit_binom <- stan_glm(cbind(Hits, NotHits ) ~ PriPos - 1, 
              family = binomial(link = "probit"), data = batting, refresh = 0)
print(probit_binom)
plot(probit_binom)
```

\newpage

So what do the coefficients mean in this model?

\vfill

```{r}
log_binom$coefficients
probit_binom$coefficients
```

\vfill

```{r}
invlogit(log_binom$coefficients) * 1000
pnorm(probit_binom$coefficients) * 1000
```




